Equivalent Income Outcome Visualisation

### 2023-09-17 minos/outcomes/EI_PresPlots_Refined.Rmd
Lots of testing on S7 component models. Also added a script to check the proportional odds assumption for deciding between clm and nnet. Still working on a function to add to the transition estimation scripts to produce a file with each model

---
title: "Equivalent Income Stuff - Energy Crisis Plots"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

@@

```{r}
S7.var.list <- c('hh_income', 'equivalent_income',  # Income variables
                 'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness',  # SIPHER 7 variables
                 'ethnicity', 'age', 'region', 'job_sec', 'education_state', 'nkids_ind', 'housing_tenure', 'urban')
out.path.batch <- here::here('output', 'SIPHER7_batch2/')

@@

@@

  ggplot(combined, aes(x = time, y = mean_diff_hh_income, group = .data[[group.var]], color = .data[[group.var]])) +
    geom_smooth() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(title = paste0('Household Income change by ', group.var), subtitle = int.name) +
    xlab('Year') +
    ylab('Household Income (£s)')
}
```

```{r}
income_equivinc_percent_change_batch <- function(base.batch, int.batch, int.name, group.var) {
  base.inc <- base.batch %>%
    select(pidp, time, run_id, hh_income, equivalent_income, all_of(group.var), weight) %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    summarise(base_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
              base_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

@@

income_equivinc_percent_change_batch(base.batch, int.batch, int.name, group.var = 'income.quint')

```{r}
## ATTEMPT 3

@@

  ggplot(data = combined, aes(x = time, y = prop_diff, group = .data[[var]], color = .data[[var]])) +
    geom_smooth() +
    facet_wrap(~ .data[[group.var]], nrow=1) +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
}
```

```{r}
S7_cumulative_pathwayChange_byGroup <- function(base.batch, int.batch, group.var, var) {
  base.sub <- base.batch %>%
    select(pidp, time, run_id, equivalent_income, all_of(group.var), all_of(var)) %>%
    group_by(time, run_id, .data[[group.var]], .data[[var]]) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    mutate(total = sum(n),
           base_prop = n / total) %>%
    select(-n, -total)

@@

  ggplot(data = combined, aes(x = .data[[var]], y = cumulative_diff, fill = .data[[var]])) +
    geom_col() +
    facet_wrap(~ .data[[group.var]], nrow=1) +
    labs(title = paste0('Cumulative Difference in ', var, ' by ', group.var)) +
    ylab('Cumulative Difference (%)') +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = 'bottom')
}
```

## Income Quintile

## Age Group

```{r}
simplified_component_change <- function(base.batch, int.batch, group.var, var, print = FALSE) {

@@

  if (var == 'S7_housing_quality') {
    base.batch$S7_housing_quality <- as.numeric(factor(base.batch$S7_housing_quality,
                                            levels = c('No to all',
                                                       'Yes to some',
                                                       'Yes to all'),
                                            labels = c(1, 2, 3)))
    int.batch$S7_housing_quality <- as.numeric(factor(int.batch$S7_housing_quality,
                                            levels = c('No to all',
                                                       'Yes to some',
                                                       'Yes to all'),
                                            labels = c(1, 2, 3)))
  } else if (var == 'S7_neighbourhood_safety') {
    base.batch$S7_neighbourhood_safety <- as.numeric(factor(base.batch$S7_neighbourhood_safety,
                                                 levels = c('Often',
                                                            'Some of the time',
                                                            'Hardly ever'),
                                                 labels = c(1, 2, 3)))
    int.batch$S7_neighbourhood_safety <- as.numeric(factor(int.batch$S7_neighbourhood_safety,
                                                 levels = c('Often',
                                                            'Some of the time',
                                                            'Hardly ever'),
                                                 labels = c(1, 2, 3)))
  } else if (var == 'loneliness') {
    base.batch$loneliness <- as.numeric(factor(base.batch$loneliness,
                                    levels = c('Often', 'Sometimes', 'Never'),
                                    labels = c(1, 2, 3)))
    int.batch$loneliness <- as.numeric(factor(int.batch$loneliness,
                                    levels = c('Often', 'Sometimes', 'Never'),
                                    labels = c(1, 2, 3)))
  }

  base.sub <- base.batch %>%
    select(pidp, time, run_id, all_of(group.var), all_of(var)) %>%
    rename(Baseline_var = .data[[var]],
           Baseline_group.var = .data[[group.var]])

  int.sub <- int.batch %>%
    select(pidp, time, run_id, all_of(group.var), all_of(var)) %>%
    rename(Intervention_var = .data[[var]],
           Intervention_group.var = .data[[group.var]])

@@

  merged2 <- merged %>%
    mutate(varChange = Intervention_var - Baseline_var,
           effect = case_when(varChange > 0 ~ "Positive",
                              varChange < 0 ~ "Negative",
                              varChange == 0 ~ "No_Change")) %>%
    select(-Baseline_var, -Intervention_var, -Intervention_group.var, -varChange) %>%
    rename(income.quint = Baseline_group.var) %>%
    group_by(time, run_id, .data[[group.var]], effect) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    mutate(total = sum(n),
           effect_prop = n / total) %>%
    select(-n, -total) %>%
    group_by(time, .data[[group.var]], effect) %>%
    summarise(effect_prop = mean(effect_prop)) %>%
    pivot_wider(names_from = 'effect',
                values_from = 'effect_prop') %>%
    mutate(change = Positive - Negative) %>%
    select(-Positive, -Negative, -No_Change, -time) %>%
    group_by(.data[[group.var]]) %>%
    summarise(change = sum(change))

  p1 <- ggplot(data = merged2, aes(x = .data[[group.var]], y = change)) +
    geom_col() +
    labs(title = paste0('Change in ', var, ' by ', group.var)) +
    ylab('% Change')

  if(print) {
    print(p1)
  }

  return(merged2)
}

simplified_component_change_BOXPLOT <- function(base.batch, int.batch, group.var, var, print = FALSE) {

  if (var == 'S7_housing_quality') {
    base.batch$S7_housing_quality <- as.numeric(factor(base.batch$S7_housing_quality,
                                            levels = c('No to all',
                                                       'Yes to some',
                                                       'Yes to all'),
                                            labels = c(1, 2, 3)))
    int.batch$S7_housing_quality <- as.numeric(factor(int.batch$S7_housing_quality,
                                            levels = c('No to all',
                                                       'Yes to some',
                                                       'Yes to all'),
                                            labels = c(1, 2, 3)))
  } else if (var == 'S7_neighbourhood_safety') {
    base.batch$S7_neighbourhood_safety <- as.numeric(factor(base.batch$S7_neighbourhood_safety,
                                                 levels = c('Often',
                                                            'Some of the time',
                                                            'Hardly ever'),
                                                 labels = c(1, 2, 3)))
    int.batch$S7_neighbourhood_safety <- as.numeric(factor(int.batch$S7_neighbourhood_safety,
                                                 levels = c('Often',
                                                            'Some of the time',
                                                            'Hardly ever'),
                                                 labels = c(1, 2, 3)))
  } else if (var == 'loneliness') {
    base.batch$loneliness <- as.numeric(factor(base.batch$loneliness,
                                    levels = c('Often',
                                               'Sometimes',
                                               'Never'),
                                    labels = c(1, 2, 3)))
    int.batch$loneliness <- as.numeric(factor(int.batch$loneliness,
                                    levels = c('Often',
                                               'Sometimes',
                                               'Never'),
                                    labels = c(1, 2, 3)))
  }

  base.sub <- base.batch %>%
    select(pidp, time, run_id, all_of(group.var), all_of(var)) %>%
    rename(Baseline_var = .data[[var]],
           Baseline_group.var = .data[[group.var]])

  int.sub <- int.batch %>%
    select(pidp, time, run_id, all_of(group.var), all_of(var)) %>%
    rename(Intervention_var = .data[[var]],
           Intervention_group.var = .data[[group.var]])

  merged2 <- merged %>%
    mutate(varChange = Intervention_var - Baseline_var,
           effect = case_when(varChange > 0 ~ "Positive",
                              varChange < 0 ~ "Negative",
                              varChange == 0 ~ "No_Change")) %>%
    select(-Baseline_var, -Intervention_var, -Intervention_group.var, -varChange) %>%
    rename(!!group.var := Baseline_group.var) %>%
    group_by(time, run_id, .data[[group.var]], effect) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    mutate(total = sum(n),
           effect_prop = n / total) %>%
    select(-n, -total) %>%
    pivot_wider(names_from = 'effect',
                values_from = 'effect_prop') %>%
    mutate(change = Positive - Negative) %>%
    select(-Positive, -Negative, -No_Change) %>%
    group_by(run_id, .data[[group.var]]) %>%
    summarise(sum.change = sum(change))

  p1 <- ggplot(data = merged2, aes(x = .data[[group.var]], y = sum.change)) +
    geom_boxplot(notch = FALSE) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(title = paste0('Change in ', var, ' by ', group.var)) +
    ylab('% Change')

  if(print) {
    print(p1)
  }

  return(merged2)
}

```{r}
# simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality', print = TRUE)
# simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'S7_neighbourhood_safety', print = TRUE)
# simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'S7_mental_health', print = TRUE)
# simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'S7_physical_health', print = TRUE)
simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'loneliness', print = TRUE)
```

```{r}
# simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality', print = TRUE)
# simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'S7_neighbourhood_safety', print = TRUE)
# simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'S7_mental_health', print = TRUE)
# simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'S7_physical_health', print = TRUE)
simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'loneliness', print = TRUE)
```

# Change in ALL vars

```{r}
varlist <- c('S7_neighbourhood_safety', 'loneliness',
             'S7_physical_health', 'S7_mental_health')

df <- simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality')
colnames(df) <- c('Income_Quintile', 'S7_housing_quality')
print("Completed for S7_housing_quality...")
for (S7_var in varlist) {
  df2 <- simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = S7_var)
  colnames(df2) <- c('Income_Quintile', S7_var)
  df <- merge(df, df2, by = c('Income_Quintile'))
  print(paste0("Completed for ", S7_var, '...'))
}
rm(df2)

# finally pivot_longer for ggplot and facet_wrap to take over
df2 <- df %>%
  pivot_longer(cols = S7_housing_quality:S7_mental_health,
               names_to = 'Variable',
               values_to = 'Change')

ggplot(df2, aes(x = Variable, y = Change, group = Variable, colour = Variable, fill = Variable)) +
  geom_col() +
  facet_wrap(~ Income_Quintile, nrow = 1) +
  theme(legend.position = 'bottom',
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```

# BOXPLOT Change in ALL vars

```{r}
varlist <- c('S7_neighbourhood_safety', 'loneliness',
             'S7_physical_health', 'S7_mental_health')

df <- simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality')
colnames(df) <- c('run_id', 'Income_Quintile', 'S7_housing_quality')
print("Completed for S7_housing_quality...")
for (S7_var in varlist) {
  df2 <- simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = S7_var)
  colnames(df2) <- c('run_id', 'Income_Quintile', S7_var)
  df <- merge(df, df2, by = c('run_id', 'Income_Quintile'))
  print(paste0("Completed for ", S7_var, '...'))
}
rm(df2)

# finally pivot_longer for ggplot and facet_wrap to take over
df2 <- df %>%
  pivot_longer(cols = S7_housing_quality:S7_mental_health,
               names_to = 'Variable',
               values_to = 'Change')

ggplot(df2, aes(x = Variable, y = Change, group = Variable, colour = Variable, fill = Variable)) +
  geom_boxplot(notch = FALSE) +
  facet_wrap(~ Income_Quintile, nrow = 1) +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  theme(legend.position = 'bottom',
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```

@@

>>>>

---
title: "Equivalent Income Presentation Plots - FINAL"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

@@

```{r}
S7.var.list <- c('hh_income', 'equivalent_income',  # Income variables
                 'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness',  # SIPHER 7 variables
                 'ethnicity', 'age', 'region', 'job_sec', 'education_state', 'nkids_ind', 'housing_tenure', 'urban')
out.path.batch <- here::here('output', 'SIPHER7_batch/')

@@

# out.path.arc2 <- here::here('output', 'batch')
# 
# #target.pattern <- paste0('aggregated_run_id_', year, '.csv')
# base.list <- list.files(path = paste0(out.path.arc2, '/baseline'),
#                             full.names = TRUE)
# base.batch <- do.call(rbind, lapply(base.list, read.csv))
# 
# energy.list <- list.files(path = paste0(out.path.arc2, '/energy'),
#                             full.names = TRUE)
# int.batch <- do.call(rbind, lapply(energy.list, read.csv))

@@

  ggplot(combined, aes(x = time, y = mean_diff_hh_income, group = .data[[group.var]], color = .data[[group.var]])) +
    geom_smooth() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(title = paste0('Household Income change by ', group.var), subtitle = int.name) +
    xlab('Year') +
    ylab('Household Income (£s)')
}

income_equivinc_percent_change_batch <- function(base.batch, int.batch, int.name, group.var) {
  base.inc <- base.batch %>%
    select(pidp, time, run_id, hh_income, equivalent_income, all_of(group.var), weight) %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    summarise(base_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
              base_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

@@

@@

  ggplot(data = combined, aes(x = time, y = prop_diff, group = .data[[var]], color = .data[[var]])) +
    geom_smooth() +
    facet_wrap(~ .data[[group.var]], nrow=1) +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
}

S7_cumulative_pathwayChange_byGroup <- function(base.batch, int.batch, group.var, var) {
  base.sub <- base.batch %>%
    select(pidp, time, run_id, equivalent_income, all_of(group.var), all_of(var)) %>%
    group_by(time, run_id, .data[[group.var]], .data[[var]]) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    mutate(total = sum(n),
           base_prop = n / total) %>%
    select(-n, -total)

@@

  ggplot(data = combined, aes(x = .data[[var]], y = cumulative_diff, fill = .data[[var]])) +
    geom_col() +
    facet_wrap(~ .data[[group.var]], nrow=1) +
    labs(title = paste0('Cumulative Difference in ', var, ' by ', group.var)) +
    ylab('Cumulative Difference (%)') +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = 'bottom')
}

simplified_component_change_BOXPLOT <- function(base.batch, int.batch, group.var, var.list, print = FALSE) {

@@

  base.batch$S7_housing_quality <- as.numeric(factor(base.batch$S7_housing_quality,
                                          levels = c('No to all',
                                                     'Yes to some',
                                                     'Yes to all'),
                                          labels = c(1, 2, 3)))
  int.batch$S7_housing_quality <- as.numeric(factor(int.batch$S7_housing_quality,
                                          levels = c('No to all',
                                                     'Yes to some',
                                                     'Yes to all'),
                                          labels = c(1, 2, 3)))
  base.batch$S7_neighbourhood_safety <- as.numeric(factor(base.batch$S7_neighbourhood_safety,
                                               levels = c('Often',
                                                          'Some of the time',
                                                          'Hardly ever'),
                                               labels = c(1, 2, 3)))
  int.batch$S7_neighbourhood_safety <- as.numeric(factor(int.batch$S7_neighbourhood_safety,
                                               levels = c('Often',
                                                          'Some of the time',
                                                          'Hardly ever'),
                                               labels = c(1, 2, 3)))
  base.batch$loneliness <- as.numeric(factor(base.batch$loneliness,
                                  levels = c('Often',
                                             'Sometimes',
                                             'Never'),
                                  labels = c(1, 2, 3)))
  int.batch$loneliness <- as.numeric(factor(int.batch$loneliness,
                                  levels = c('Often',
                                             'Sometimes',
                                             'Never'),
                                  labels = c(1, 2, 3)))

  base.sub <- base.batch %>%
    select(pidp, time, run_id, all_of(group.var), all_of(var.list)) %>%
    rename_with(.fn = ~paste0('Baseline_', .), .cols = all_of(var.list))

  int.sub <- int.batch %>%
    select(pidp, time, run_id, all_of(var.list)) %>%
    rename_with(.fn = ~paste0('Intervention_', .), .cols = all_of(var.list))

@@

  # https://stackoverflow.com/questions/49694353/dplyr-mutating-multiple-columns-by-prefix-and-suffix

  merged2 <- map(var.list, ~ merged %>%
                   mutate(!!paste0("Change_", .x) := !!as.name(paste0("Intervention_", .x)) - !!as.name(paste0("Baseline_", .x)))) %>%
    reduce(left_join) %>%
    select(-starts_with('Baseline_'), -starts_with('Intervention_')) %>%
    pivot_longer(cols = paste0('Change_', first(var.list)):paste0('Change_', last(var.list)),
                 names_to = 'Variable',
                 names_prefix = 'Change_',
                 values_to = 'Change') %>%
    mutate(effect = case_when(Change > 0 ~ "Positive",
                              Change < 0 ~ "Negative",
                              Change == 0 ~ "No_Change")) %>%
    select(-Change) %>%
    group_by(time, run_id, .data[[group.var]], Variable, effect) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]], Variable) %>%
    mutate(total = sum(n),
           effect_prop = (n / total) * 100) %>%
    select(-n, -total) %>%
    pivot_wider(names_from = 'effect',
                values_from = 'effect_prop') %>%
    mutate(change = Positive - Negative) %>%
    select(-Positive, -Negative, -No_Change) %>%
    group_by(run_id, .data[[group.var]], Variable) %>%
    summarise(sum.change = sum(change))

  ggplot(merged2, aes(x = Variable, y = sum.change, group = Variable, colour = Variable, fill = Variable)) +
    geom_boxplot(notch = FALSE) +
    facet_wrap(~ .data[[group.var]], nrow = 1) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(title = paste0("Change in component variables by ", group.var)) +
    ylab("Net Change (%)") +
    theme(legend.position = 'bottom',
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank())

}
```

## Absolute Change

@@

```{r}
income_equivinc_percent_change_batch(base.batch, int.batch, int.name, group.var = 'income.quint')
```

### Income Quintile

### Age Group

# Simplified Cumulative Change

```{r}
varlist <- c('S7_housing_quality', 'S7_neighbourhood_safety', 'loneliness',
             'S7_physical_health', 'S7_mental_health')

simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var.list = varlist)
```

### 2023-09-17 minos/outcomes/equivalent_income_out.Rmd
Lots of testing on S7 component models. Also added a script to check the proportional odds assumption for deciding between clm and nnet. Still working on a function to add to the transition estimation scripts to produce a file with each model

```{r}
# set output path to SIPHER 7 experiment out
out.path <- here::here('output', 'SIPHER7/')
# read all files into dataframe (util function handles runtime subdirectory etc.)
base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
# allchild.dat <- read_singular_local_out(out.path, 'hhIncomeChildUplift', drop.dead = TRUE)
# pov.dat <- read_singular_local_out(out.path, 'hhIncomePovertyLineChildUplift', drop.dead = TRUE)
# wage.dat <- read_singular_local_out(out.path, 'livingWageIntervention', drop.dead = TRUE)
energy.dat <- read_singular_local_out(out.path, 'energyDownlift', drop.dead = TRUE)
```

>>>>

```{r}
# set output path to SIPHER 7 experiment out
out.path <- here::here('output', 'SIPHER7asd/')
# read all files into dataframe (util function handles runtime subdirectory etc.)
base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
# allchild.dat <- read_singular_local_out(out.path, 'hhIncomeChildUplift', drop.dead = TRUE)
# pov.dat <- read_singular_local_out(out.path, 'hhIncomePovertyLineChildUplift', drop.dead = TRUE)
# wage.dat <- read_singular_local_out(out.path, 'livingWageIntervention', drop.dead = TRUE)
energy.dat <- read_singular_local_out(out.path, 'energyDownlift', drop.dead = TRUE)
```
